### 2014-03-01 paper/related.tex
I don't think we can consider langs with non-derministic relations, in general
Related work, pass

Our paper follows a long tradition of papers on information flow
control.  Our information flow control system is mostly closely related
to the coarse-grained information systems used in operating systems such
as Asbestos~\cite{efstathopoulos:asbestos} and
HiStar~\cite{Zeldovich:2006}, as well as language-level
\emph{floating-label IFC systems} such as LIO~\cite{lio},
Breeze~\cite{Hritcu:2013:YIB:2497621.2498098}, where there is a label
associated with threads of execution which increases monotonically.
Our treatment of termination-sensitive and termination-insensitive interference
originates from Smith and Volpano~\cite{Smith:Volpano:MultiThreaded,Volpano:1997:ECF:794197.795081}.
While this work often considers the interaction of information flow
control with respect to various language features (e.g.,
exceptions~\cite{Hritcu:2013:YIB:2497621.2498098}), none of these
systems attempt to be generic in the choice of target language.

One of the primary motivations behind this paper is constructing a
practical information flow control system for JavaScript.  Some of
these systems attempt to perform fine-grained information flow of
JavaScript, e.g. Hedin and Sabelfield's work~\cite{Hedin:2012} on JSFlow~\cite{JSFlow}
and \Red{some other stuff here}.  While fine-grained information flow
control can result in very precise analysis of legacy code, it comes
at the cost of high theoretical complexity (the system must accomodate
the entirety of JavaScript's semantics) and runtime performance cost.
There are attempts to overcome this performance barrier by various means:
Austin and Flanagan~\cite{Austin:Flanagan:PLAS09} handle this problem
by using a \emph{sparse information labeling} scheme which reduces the
cost of managing labels when they are not needed, while FlowFox~\cite{DeGroef:2012}
uses an entirely different technique of \emph{secure multi-execution} to
enforce non-interference. \Red{Probably more stuff here.}

The constructs in our IFC language, as well as the behavior of
inter-thread communication, is reminiscent of distributed systems
like Erlang.~\cite{Armstrong03makingreliable}  There is a close
correspondence: in distributed systems, isolation is required due to
physical constraints; in information flow control, isolation is
required to enforce noninterference.  Papagiannis et al.~\cite{Papagiannis_enforcinguser}
built an information flow control system on top of Erlang which is quite
similar to ours.  However, they do not use a floating label system (processes
can find out when sending a message failed due to a forbidden information flow),
and they include no security proofs.

There appears to be limited literature in general techniques for
retrofitting information flow control to arbitrary languages; however,
one time-honored technique is defining a fundamental calculus, for which
other languages can be desugared into.  Abadi et al.~\cite{abadi+:core}
motivate their core calculus of dependency by showing how various
previous systems can be encoded in it; Tse and
Zdancewic~\cite{Tse:Zdancewic:ICFP04}, in turn, show how this calculus
can be encoded in System F via parametricity.  However, this work is
primarily focused on static information flow control systems.

@@

It has long been observed that information flow control and monads are
closely related.  This observation has been used by Abadi et
al.~\cite{abadi+:core} to structure access to private data by protecting
it using a lattice of monads, one per security level.  Our system uses
monads in a conceptually different manner, where there is a single monad
representing the information flow control computational effects; this
is inline with the use of monads in other systems.~\cite{Harrison05,lio,Devriese:2011}.
Devriese and Piessens~\cite{Devriese:2011} utilized monad
transformers as an essential component in their work.  However, their use
of monad transformers is to explain when extra constraints on a non-information
flow control aware base monad should be applied (i.e., during lifting).
Harrison and Hook~\cite{Harrison05} define a notion of \emph{atomic noninterference},
which is preserved through monad transformer application.  However, it is not
clear what the relationship between this notion and traditional noninterference is,
and in their paper they prove only a weaker property of no write down.

>>>>

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End:

%Our paper follows a long tradition of papers on information flow
%control.  
Our information flow control system is closely related
to the coarse-grained information systems used in operating systems such
as Asbestos~\cite{efstathopoulos:asbestos}, 
HiStar~\cite{Zeldovich:2006}, and Flume~\cite{krohn:flume}, as well as language-based
\emph{floating-label IFC systems} such as LIO~\cite{lio},
and Breeze~\cite{Hritcu:2013:YIB:2497621.2498098}, where there is a
monotonically increased label
associated with threads of execution.
Our treatment of termination-sensitive and termination-insensitive interference
originates from Smith and Volpano~\cite{Smith:Volpano:MultiThreaded,Volpano:1997:ECF:794197.795081}.
%While this work often considers the interaction of information flow
%control with respect to various language features (e.g.,
%exceptions~\cite{Hritcu:2013:YIB:2497621.2498098}), none of these
%systems attempt to be generic in the choice of target language.

Designed to target legacy code, secure multi-execution (SME) is a black-box
technique to enforce information-flow control~\cite{Devriese:2010}. It runs
multiple copies of the program, one per security level, where the semantics of
I/O interactions is altered. Besides its black-box nature, authors often present
it considering one specific programming
language~\cite{KULeuven-350547,Rafnson:2013}. Differently, Bielova
al.~\cite{Biel-etal-11-TR} uses a transition system to describe SME, where the
details of the underlying language is hidden.  Zanarini et
al.~\cite{ZanariniJR13} propose a novel semantics for programs based on
interaction trees~\cite{jacobs-tutorial}, which threats programs as black-boxes
about which nothing is known except what is inferred from their I/O interactions
with the environment. Similar to SME, our approach only mediates some I/O
operations; differently, it runs the program once.

One of the primary motivations behind this paper is constructing a practical
information flow control system for JavaScript.  Some of these systems attempt
to perform fine-grained information flow of
JavaScript~\cite{Hedin:2012,ConDOM,JSFlow}. While fine-grained information flow
control can result in less false alarms and target legacy code, it comes at the
cost of complexity (the system must accomodate the entirety of JavaScript's
semantics). Moreover, the runtime performance cost are sometimes too high, e.g.,
higher than 100\%!\cite{JSFlow}.  Alternatively, coarse-grained approaches have
also been proposed for the web
scenario~\cite{Yip:2009:PBS,DeGroef:2012,conf/esorics/AkhaweLHSS13}. Different
from them, we rely on programming level abstraction to isolate
computations.

%There are attempts to overcome this performance barrier by various means:
%Austin and Flanagan~\cite{Austin:Flanagan:PLAS09} handle this problem
%by using a \emph{sparse information labeling} scheme which reduces the
%cost of managing labels when they are not needed, while FlowFox~\cite{DeGroef:2012}
%uses an entirely different technique of \emph{secure multi-execution} to
%enforce non-interference. \Red{Probably more stuff here.}
%\todo{ds}{SME works for arbitrary language (kind of).}

The constructs in our IFC language, as well as the behavior of
inter-thread communication, is reminiscent of distributed systems
like Erlang~\cite{Armstrong03makingreliable}.  There is a close
correspondence: in distributed systems, isolation is required due to
physical constraints; in information flow control, isolation is
required to enforce noninterference.  Papagiannis et al.~\cite{Papagiannis_enforcinguser}
built an information flow control system on top of Erlang which is quite
similar to ours.  However, they do not use a floating label system (processes
can find out when sending a message failed due to a forbidden information flow),
and they include no security proofs.

There appears to be limited literature in general techniques for retrofitting
information flow control to arbitrary languages; however, one time-honored
technique is defining a fundamental calculus, for which other languages can be
desugared into.  Abadi et al.~\cite{abadi+:core} motivate their core calculus of
dependency by showing how various previous systems can be encoded in it; Tse and
Zdancewic~\cite{Tse:Zdancewic:ICFP04}, in turn, show how this calculus can be
encoded in System F via parametricity.  Broberg and Sands~\cite{Broberg:2010}
encodes several IFC systems into Paralocks.  However, this line of work is
primarily focused on static enforcements. 
%information flow control systems.

@@

It has long been observed that information flow control and monads are
closely related.  This observation has been used by Abadi et
al.~\cite{abadi+:core} to structure access to private data by protecting
it using a lattice of monads, one per security level. Russo et. al.~\cite{Russo+:Haskell08} extend this 
idea to languages with side-effects. Our system uses
monads in a conceptually different manner, where there is a single monad
representing the information flow control computational effects; this
is inline with the use of monads in other systems~\cite{Harrison05,lio,Devriese:2011}.
Devriese and Piessens~\cite{Devriese:2011} utilize monad
transformers as an essential component in their work.  However, their use
of monad transformers is to explain when extra constraints on a non-information
flow control aware base monad should be applied (i.e., during lifting).
Harrison and Hook~\cite{Harrison05} define a notion of \emph{atomic noninterference},
which is preserved through monad transformer application.  However, it is not
clear what is the relationship between this notion and traditional noninterference,
and their paper only proves a weaker property of no write down.
